---
title: "gDRplots"
author: "gDR team"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{gDRtestData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r lib, echo=FALSE}
library(gDRplots)
```

# Overview

The `gDRplots` package belongs to the `gDR` suite apps. 

# Content

The `gDRplots` package is intended to store functions for gDR static visualizations and their dedicated helpers. The package allows users to visualize different metrics for combinations of drugs and cell lines of interest - during and after processing.

The `gDRplots` package supports metrics: 

```{r metrics_def, echo = FALSE}
ls_metrics <- c("GR value", "Relative Viability")
dict <- gDRcomponents::get_dict()
metrics_def <- data.table::data.table(
  Metrics = sprintf("**%s**", ls_metrics),
  Definition = vapply(ls_metrics, function(i) dict[[i]], character(1))
)
knitr::kable(metrics_def, escape = FALSE)
```

Note, all plot function required as an input data.table representation of the data in selected assaye with added information from colData created based on the gDR data model (MultiAssayExperiment).

## Dose Response Plots

The Dose Response Plots is classic drug response curves visualization in a consolidated way. 

The overview with many curves presented on the same scale gives the user a big picture of the dataset, especially in a big compound screening dataset, in which the drug has a universal resistant pattern across all cell lines and becomes easily spotted. The compound generated different sigmoid curves suggesting inhibitory or cytotoxic effects which is worth exploring further. 

### Single-agent Data

The **plot_dose_response_sa** function is dedicated for combination experiments.

User can also use  **plot_dose_response_sa_by_CLs** or **plot_dose_response_sa_by_drugs** to get list of dose response plots by cell line names or by drugs accordingly.

All above functions required as an input two tables representing _the metrics assay_ and _the average assay_.

```{r plot_dose_response_sa, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("small")
se <- mae[[gDRutils::get_supported_experiments("sa")]]

dt_metrics <- gDRutils::convert_se_assay_to_dt(
  se["G00002_drug_002_moa_A_72"], assay_name = "Metrics")
dt_average <- gDRutils::convert_se_assay_to_dt(
  se["G00002_drug_002_moa_A_72"], assay_name = "Averaged")

plot_dose_response_sa(
  dt_metrics = dt_metrics,
  dt_average = dt_average,
  grouping = "CellLineName",
  group_names = c("cellline_BA", "cellline_CA", "cellline_DA"))
```

### Combination Data

The **plot_dose_response_combo** function is dedicated for combination experiments.

User can also use  **plot_dose_response_combo_qc_panel** to get panel of dose response plots by selected cell line name..

This function required as an input a table representing _the average assay_ and cell line name, and selected drug names and co-drug name.

```{r plot_dose_response_combo, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("combo_matrix")
se <- mae[[gDRutils::get_supported_experiments("combo")]]

dt_average <- 
  gDRutils::convert_se_assay_to_dt(se, 
                                   assay_name = "Averaged")

plot_dose_response_combo(dt_average = dt_average,
                         drug1_name = "drug_011",
                         drug2_name = "drug_021",
                         cl_name = "cellline_BC")
```

## Metrics plot

### Single-agent Data

**pheatmap_with_anno_sa**

```{r pheatmap_with_anno_sa, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("combo_matrix")
se <- mae[[gDRutils::get_supported_experiments("sa")]]
response_metrics <- 
  gDRutils::convert_se_assay_to_dt(se = se,
                                   assay_name = "Metrics")

annotation_manual_col <-
  unique(response_metrics[,.SD, .SDcols = c("CellLineName", "Tissue")])
annotation_manual_row <-
  unique(response_metrics[,.SD, .SDcols = c("DrugName", "drug_moa")])

output <- pheatmap_with_anno_sa(tab_response = response_metrics,
                                metric = "x_mean",
                                annotation_row = annotation_manual_row,
                                annotation_col = annotation_manual_col)
hm <- output[["heatmap"]]
ggpubr::as_ggplot(hm[["gtable"]])

knitr::kable(output[["data"]][["matrix"]], escape = FALSE)
```

### Combination Data

**pheatmap_with_anno_combo**

**heatmap_combo_metrics**

**heatmap_combo_with_isoref**

## Quality Control 

### Quality controls of mapping controls to treated

SA and combo

**heatmap_control_mapping_qc**

### Quality Control of degree of errors in normalization

SA nad combo

**plot_var_distribution_qc**


### Quality Control for Averaged 

SA and combo 

**pheatmap_qc**


### Quality Control - Accuarcy of Fitting Dose-response Curves

Only SA

**plot_dose_response_sa_qc**
**plot_dose_response_sa_qc_panel**


### Quality Control - Fitting Precision 

Only SA

**plot_fitting_acc**

```{r plot_fitting_acc, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("combo_matrix")
se <- mae[[gDRutils::get_supported_experiments("sa")]]

dt_metrics <- 
  gDRutils::convert_se_assay_to_dt(se, 
                                   assay_name = "Metrics")

plot_fitting_acc(dt_assay = dt_metrics,
                 cl_name = "cellline_NE")
```

# SessionInfo {-}

```{r sessionInfo}
sessionInfo()
```
