---
title: "gDRplots"
author: "gDR team"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{gDRtestData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r lib, echo=FALSE}
library(gDRplots)
```

# Overview

The `gDRplots` package belongs to the `gDR` suite apps. 

# Content

The `gDRplots` package is intended to store functions for gDR static visualizations and their dedicated helpers. The package allows users to visualize different metrics for combinations of drugs and cell lines of interest - during and after processing.

The `gDRplots` package supports metrics: 

|: Metrics |: Definition |
| ---- | ---------- |
| **GR value** | Normalized Growth Rate (GR) value; calculated by normalizing treated samples with the corresponding vehicle-treated samples. |
| **Relative Viability** | (RV); calculated by normalizing treated samples with the corresponding vehicle-treated samples. |

Note, all plot function required as an input data.table representation of the data in selected assaye with added information from colData created based on the gDR data model (MultiAssayExperiment).

Used data:

```{r sa_data, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("combo_matrix")
se_sa <- mae[[gDRutils::get_supported_experiments("sa")]]

dt_norm_sa <- 
  gDRutils::convert_se_assay_to_dt(se = se_sa,
                                   assay_name =  "Normalized")
dt_metrics_sa <- 
  gDRutils::convert_se_assay_to_dt(se = se_sa,
                                   assay_name = "Metrics")
dt_average_sa <- 
  gDRutils::convert_se_assay_to_dt(se = se_sa,
                                   assay_name = "Averaged")
```

```{r combo_data, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("combo_matrix")
se_combo <- mae[[gDRutils::get_supported_experiments("combo")]]

dt_norm_combo <- 
  gDRutils::convert_se_assay_to_dt(se = se_combo,
                                   assay_name =  "Normalized")
dt_metrics_combo <- 
  gDRutils::convert_se_assay_to_dt(se = se_combo, 
                                   assay_name = "Metrics")
dt_average_combo <- 
  gDRutils::convert_se_assay_to_dt(se = se_combo, 
                                   assay_name = "Averaged")
dt_scores <- 
  gDRutils::convert_se_assay_to_dt(se = se_combo,
                                   assay_name = "scores")
dt_excess <- 
  gDRutils::convert_se_assay_to_dt(se = se_combo,
                                   assay_name = "excess")
dt_isobolograms <- 
  gDRutils::convert_se_assay_to_dt(se = se_combo,
                                   assay_name = "isobolograms")
```

## Dose Response Plots

The Dose Response Plots is classic drug response curves visualization in a consolidated way. 

The overview with many curves presented on the same scale gives the user a big picture of the dataset, especially in a big compound screening dataset, in which the drug has a universal resistant pattern across all cell lines and becomes easily spotted. The compound generated different sigmoid curves suggesting inhibitory or cytotoxic effects which is worth exploring further. 

### Single-agent Data

The **plot_dose_response_sa** function is dedicated for combination experiments.

User can also use  **plot_dose_response_sa_by_CLs** or **plot_dose_response_sa_by_drugs** to get list of dose response plots by cell line names or by drugs accordingly.

All above functions required as an input two tables representing _the metrics assay_ and _the average assay_.

```{r plot_dose_response_sa, warning = FALSE, message = FALSE}
mae <- gDRutils::get_synthetic_data("small")
se <- mae[[gDRutils::get_supported_experiments("sa")]]

dt_metrics <- gDRutils::convert_se_assay_to_dt(
  se["G00002_drug_002_moa_A_72"], assay_name = "Metrics")
dt_average <- gDRutils::convert_se_assay_to_dt(
  se["G00002_drug_002_moa_A_72"], assay_name = "Averaged")

plot_dose_response_sa(
  dt_metrics = dt_metrics,
  dt_average = dt_average,
  grouping = "CellLineName",
  group_names = c("cellline_BA", "cellline_CA", "cellline_DA"))
```

### Combination Data

The **plot_dose_response_combo** function is dedicated for combination experiments.

User can also use **plot_dose_response_combo_panel** to get panel of dose response plots by selected cell line name.

This function required as an input a table representing _the average assay_ and cell line name, and selected drug names and co-drug name.

```{r plot_dose_response_combo, warning = FALSE, message = FALSE}
plot_dose_response_combo(dt_average = dt_average_combo,
                         drug1_name = "drug_011",
                         drug2_name = "drug_021",
                         cl_name = "cellline_BC")
```

## Metrics plot

### Single-agent Data

**pheatmap_with_anno_sa**

```{r pheatmap_with_anno_sa, warning = FALSE, message = FALSE}
annotation_manual_col <-
  unique(dt_metrics_sa[, .SD, .SDcols = c("CellLineName", "Tissue")])
annotation_manual_row <-
  unique(dt_metrics_sa[, .SD, .SDcols = c("DrugName", "drug_moa")])

output <- pheatmap_with_anno_sa(
  dt_metrics = dt_metrics_sa,
  annotation_row = annotation_manual_row,
  annotation_col = annotation_manual_col)
hm <- output[["heatmap"]]
ggpubr::as_ggplot(hm[["gtable"]])

knitr::kable(output[["data"]][["matrix"]], escape = FALSE)
```

### Combination Data

For experiments with drug combinations, user can visualize the different combination metrics in a summary heatmaps where 

**pheatmap_with_anno_combo**

```{r pheatmap_with_anno_combo, warning = FALSE, message = FALSE}
annotation_manual_col <-
  unique(dt_scores[,.SD, .SDcols = c("CellLineName", "Tissue")])
annotation_manual_row <-
  unique(dt_scores[,.SD, .SDcols = c("DrugName", "DrugName_2", "drug_moa", "drug_moa_2")])

output <- pheatmap_with_anno_combo(
  dt_scores = dt_scores,
  annotation_row = annotation_manual_row,
  annotation_col = annotation_manual_col)
hm <- output[["heatmap"]]
ggpubr::as_ggplot(hm[["gtable"]])

knitr::kable(output[["data"]][["matrix"]], escape = FALSE)
```

**heatmap_combo_metrics**

```{r heatmap_combo_metrics, warning = FALSE, message = FALSE}
heatmap_combo_metrics(dt_excess = dt_excess,
                      dt_isobolograms = dt_isobolograms,
                      drug1_name = "drug_001", 
                      drug2_name = "drug_026",
                      cl_name = "cellline_BC",
                      iso_levels = "0.5")
```

**heatmap_combo_with_isoref**

User can also use **heatmap_combo_with_isoref_panel** to get panel with heatmap for selected drug namea nad co-drug name and cell line names list.

```{r heatmap_combo_with_isoref, warning = FALSE, message = FALSE}
heatmap_combo_with_isoref(dt_excess = dt_excess,
                          dt_isobolograms = dt_isobolograms,
                          drug1_name = "drug_001", 
                          drug2_name = "drug_026",
                          cl_name = "cellline_BC",
                          iso_levels = c("0.2", "0.5", "0.8"))
```


## Quality Control 

### Mapping Controls to Treated

SA and combo

**heatmap_control_mapping_qc**

```{r heatmap_control_mapping_qc, warning = FALSE, message = FALSE}
dt_treat <- 
  gDRutils::convert_se_assay_to_dt(se_sa, 
                                   assay_name = "RawTreated")
dt_controls <- 
  gDRutils::convert_se_assay_to_dt(se_sa, 
                                   assay_name = "Controls")

heatmap_control_mapping_qc(dt_treat = dt_treat,
                           dt_controls = dt_controls)


dt_treat <- 
  gDRutils::convert_se_assay_to_dt(se_combo, 
                                   assay_name = "RawTreated")
dt_controls <- 
  gDRutils::convert_se_assay_to_dt(se_combo, 
                                   assay_name = "Controls")

heatmap_control_mapping_qc(dt_treat = dt_treat,
                           dt_controls = dt_controls)
```

### Degree of Errors in Normalization Values

SA nad combo

**plot_var_distribution_qc**

```{r plot_var_distribution_qc, warning = FALSE, message = FALSE}
plot_var_distribution_qc(dt_assay = dt_norm_sa,
                         cl_name = "cellline_BC")

plot_var_distribution_qc(dt_assay = dt_norm_combo,
                         cl_name = "cellline_BC")
```

### Averaged Values

SA and combo 

**pheatmap_qc**

```{r pheatmap_qc_sa, warning = FALSE, message = FALSE}
hm_sa <- pheatmap_qc(dt_average = dt_average_sa)
ggpubr::as_ggplot(hm_sa[["gtable"]])
```

```{r pheatmap_qc_combo, warning = FALSE, message = FALSE}
hm_combo <- pheatmap_qc(dt_average = dt_average_combo)
ggpubr::as_ggplot(hm_combo[["gtable"]])
```

### Accuarcy of Fitting Dose-response Curves

Only SA

**plot_dose_response_sa_qc**

User can also use **plot_dose_response_sa_qc_panel** to get panel with heatmap for selected drug namea nad co-drug name and cell line names list.

```{r plot_dose_response_sa_qc, warning = FALSE, message = FALSE}
plot_dose_response_sa_qc(dt_metrics = dt_metrics_sa,
                         dt_average = dt_average_sa,
                         cl_name = "cellline_BC",
                         d_name = "drug_001")
```
### Quality Control - Fitting Precision 

Only SA

**plot_fitting_acc**


```{r plot_fitting_acc, warning = FALSE, message = FALSE}
plot_fitting_acc(dt_assay = dt_metrics_sa,
                 cl_name = "cellline_NE")
```

# SessionInfo {-}

```{r sessionInfo}
sessionInfo()
```
